Thesis: draft

### 2021-04-20 draft.tex
latest update

  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution. Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?

@@

  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $50/50$ opinion distribution]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$. The distribution is normalized by dividing for the same quantity computed at $t=0$. \textbf{(a)} Results for the REC method; \textbf{(b)} Results for the PR method. Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: The Erd\H{o}s-R\'{e}nyi model (ER) with $p = 0.01$; The stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; The stochastic Watts-Strogatz model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; The Watts-Strogatz model (WS) with $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$). The starting conditions are $P_0(0) = 0.5$ and $P_1(0) = 0.5$. Each plot is the average of $10^2$ independent simulations.}

@@

  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution. Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_1000-6_PR_20-80_8x7.png}
    \caption{WS, $N = 10^3,\ K=6,\ \beta=0.01$}
    \label{ws1000}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10000-6_PR_20-80_8x7.png}
    \caption{WS, $N = 10^4,\ K=6,\ \beta=0.01$}
    \label{ws10000}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution for a WS model with $N=10^3$ or $N=10^4$, $K=6$ and $\beta = 0.01$, PR]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution. The Watts-Strogatz network is depicted with \textbf{(a)} $N=10^3,\ K=6,\ \beta=0.01$ and \textbf{(b)} $N=10^4,\ K=6,\ \beta=0.01$. In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the PR filtering method is used. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
\label{ev_op_WS_k=6}
\end{figure}
\newpage
\noindent
\textbf{Echo chambers}\\

@@

  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$. The distribution is normalized by dividing for the same quantity computed at $t=0$. Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm. The starting conditions are $P_0(0) = 0.2$ and $P_1(0) = 0.8$. Each plot is the average of $10^2$ independent simulations.}

@@

\begin{figure}[H]
	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_modularity_edge_prob_same_deg_PR_REC_12x10}
	\captionsetup{format=plain}
	\caption[Relation between the formation of echo chambers and community structure (modularity)]{The relation between the formation of echo chambers and the community structure, represented by the modularity $Q$, for both the PR and REC filtering algorithms. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). The initial opinion distribution is $50/50$. The data point with $Q = 0$ is the result for an ER network with $N = 10^3$, $p = 0.01$ and an average degree around 10. The other four data points are the results for SBM's with 10 communities of size 100 (total number of nodes $N = 10^3$), each with a different modularity (the modularity is changed by changing the edge probabilities) and an average degree around $\left<k\right> \sim 10$. Each data point is the result of $10^2$ independent simulations.}
\label{echo_vs_mod}
\end{figure}

@@

	\caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR filtering algorithm and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}

@@

  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution. \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_REC_8x7.png}
    \caption{ER}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_REC_8x7.png}
    \caption{WS}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_REC_8x7.png}
    \caption{SBM}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_REC_8x7.png}
    \caption{SBM-WS}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$). Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
\label{ev_op_50_50_stubb_REC}
\end{figure}

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_8x7.png}
    \caption{ER}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_8x7.png}
    \caption{WS}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_8x7.png}
    \caption{SBM}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_8x7.png}
    \caption{SBM-WS}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$)]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$). Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
\label{ev_op_50_50_stubb_PR}
\end{figure}
\noindent
Figure \ref{ev_op_50_50_stubb_PR} shows the results for the evolution of the fraction of opinion 0 in the system for an initial $50/50$ opinion distribution and when using the PR filtering algorithm. Data points are again represented together with the standard deviation $\sigma$. Three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$) are compared to each other for the four different network models. The results don't differ from those obtained for the REC filtering algorithm. The status-quo is maintained, the initial $50/50$ opinion distribution does not change over time when introducing stubborn people and using the PR filtering algorithm.\\
\newline
Up till now, we have discussed the results for the formation of echo chambers and the evolution of the opinions when introducing a fraction of completely stubborn nodes in the system for an initial $50/50$ opinion distribution. Lets now look at the results for introducing a fraction of stubborn agents when we are dealing with an unbalanced initial opinion distribution, eg. an initial $20/80$ distribution.\\
\newline

>>>>

  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution. Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS), $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$); \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?

@@

  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $50/50$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$. The distribution is normalized by dividing for the same quantity computed at $t=0$. \textbf{(a)} Results for the REC method; \textbf{(b)} Results for the PR method. Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: The Erd\H{o}s-R\'{e}nyi model (ER) with $p = 0.01$; The stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; The stochastic Watts-Strogatz model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; The Watts-Strogatz model (WS) with $K = 10$, two slightly different values of the clustering coefficient are used ($\left<C\right> \sim 0.55$ and $\left<C\right> \sim 0.60$). The starting conditions are $P_0(0) = 0.5$ and $P_1(0) = 0.5$. Each plot is the average of $10^2$ independent simulations.}

@@

  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution. Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_1000-6_PR_20-80_8x7.png}
    \caption{WS, $N = 10^3,\ K=6,\ \beta=0.01$}
    \label{ws1000}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_10000-6_PR_20-80_8x7.png}
    \caption{WS, $N = 10^4,\ K=6,\ \beta=0.01$}
    \label{ws10000}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ opinion distribution for a WS model with $N=10^3$ or $N=10^4$, $K=6$ and $\beta = 0.01$, PR.]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $20/80$ distribution. The Watts-Strogatz network is depicted with \textbf{(a)} $N=10^3,\ K=6,\ \beta=0.01$ and \textbf{(b)} $N=10^4,\ K=6,\ \beta=0.01$. In both cases the average clustering coefficient is around $\left<C\right> \sim 0.58$ and the PR filtering method is used. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}
\label{ev_op_WS_k=6}
\end{figure}
\newpage
\noindent
\textbf{Echo chambers}\\

@@

  \caption[Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t=500$, $\left<P_1^{\text{nn}}\right>$, for an initial $20/80$ opinion distribution.]{Distribution of the fraction of friends (nearest neighbors, nn) of node $i$ with the same opinion 1 at $t = 500$. The distribution is normalized by dividing for the same quantity computed at $t=0$. Four different network models are used, each with a total number of nodes $N = 10^3$ and an average degree around $\left<k\right> \sim 10$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. For each model two different filtering algorithms are compared to each other: the REC and the PR filtering algorithm. The starting conditions are $P_0(0) = 0.2$ and $P_1(0) = 0.8$. Each plot is the average of $10^2$ independent simulations.}

@@

\begin{figure}[H]
	\includegraphics[width=12cm, height=10cm]{echo_chamber_vs_modularity_edge_prob_same_deg_PR_REC_12x10}
	\captionsetup{format=plain}
	\caption[Relation between the formation of echo chambers and community structure (modularity).]{The relation between the formation of echo chambers and the community structure, represented by the modularity $Q$, for both the PR and REC filtering algorithms. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). The initial opinion distribution is $50/50$. The data point with $Q = 0$ is the result for an ER network with $N = 10^3$, $p = 0.01$ and an average degree around 10. The other four data points are the results for SBM's with 10 communities of size 100 (total number of nodes $N = 10^3$), each with a different modularity (the modularity is changed by changing the edge probabilities) and an average degree around $\left<k\right> \sim 10$. Each data point is the result of $10^2$ independent simulations.}
\label{echo_vs_mod}
\end{figure}

@@

	\caption[Formation of echo chambers versus fraction of completely stubborn nodes for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution and using the PR filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}

@@

  \caption[Formation of echo chambers versus fraction of completely stubborn nodes for the REF and REC filtering algorithms and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn nodes, fracRes, for an initial $50/50$ opinion distribution. \textbf{(a)} Using the REF filtering algorithm; \textbf{(b)} Using the REC filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion 0 and the normalized fraction of nodes with all neighbors having opinion 1. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<C\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the result of $10^2$ independent simulations.}

@@

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_REC_8x7.png}
    \caption{ER}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_REC_8x7.png}
    \caption{WS}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_REC_8x7.png}
    \caption{SBM}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_REC_8x7.png}
    \caption{SBM-WS}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the REC filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$). Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
\label{ev_op_50_50_stubb_REC}
\end{figure}

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_8x7.png}
    \caption{ER}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_8x7.png}
    \caption{WS}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_8x7.png}
    \caption{SBM}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_8x7.png}
    \caption{SBM-WS}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion 0 ($P_0(t)$) over time for an initial $50/50$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn people (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$). Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<C\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<C\right> \sim 0.55$ and $Q \sim 0.9$. Each plot is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
\label{ev_op_50_50_stubb_PR}
\end{figure}
\noindent
Figure \ref{ev_op_50_50_stubb_PR} shows the results for the evolution of the fraction of opinion 0 in the system for an initial $50/50$ opinion distribution and when using the PR filtering algorithm. Data points are again represented together with the standard deviation $\sigma$. Three different fractions of stubborn people (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$) are compared to each other for the four different network models. The results don't differ from those obtained for the REC filtering algorithm. The status-quo is maintained, the initial $50/50$ opinion distribution does not change over time when introducing stubborn people and using the PR filtering algorithm.\\
\newline
Up till now, we have discussed the results for the formation of echo chambers and the evolution of the opinions when introducing a fraction of completely stubborn nodes in the system for an initial $50/50$ opinion distribution. Lets now look at the results for introducing a fraction of stubborn agents when we are dealing with an unbalanced initial opinion distribution, eg. an initial $20/80$ distribution.\\
\newline
